---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sysrevdata

<!-- badges: start -->
<!-- [![Travis build status](https://travis-ci.com/softloud/sysrevdata.svg?branch=master)](https://travis-ci.com/softloud/sysrevdata) -->
<!-- badges: end -->

The goal of sysrevdata is to provide examples and extensible toolchains for structuring data for systematic review. 

See vignettes (click on the "Articles" tab above) for example toolchain walkthroughs.

## Vignettes

- Narrative synthesis table | Condensed summaries of data for human interpretation
- Machine readable data for complex visualisations and analyses | Structuring data for machine interpretation

## Example dataset

This package currently contains `bufferstrips` dataset from a published systematic review. 

```{r}
library(sysrevdata)

# take a peek at the first 10 rows, 
# restricted to first 5 columns, 
# of each dataset

bufferstrips %>%
  # select first five columns
  dplyr::select(1:5) %>% 
  # select first 10 rows
  head(10)

```


See

- `?bufferstrips`

for more details.

## Installation

You can install 
<!-- the released version of sysrevdata from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("sysrevdata") -->
<!-- ``` -->

<!-- And  -->
the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("softloud/sysrevdata")
```

# Contributing to this 

This is work that extended from ideas generated at the [Evidence Synthesis Hackathon](https://www.eshackathon.org/); it is the voluntary work of open scientists who welcome contributions. If you have an idea for an improvement or extension, please consider contributing a pull request or opening an [issue](https://github.com/softloud/sysrevdata/issues). 
