---
title: "wide-to-condensed"
description: |
  Structuring data for humans
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r pkg, message=FALSE}
# packages used in this walkthrough
library(tidyverse) # all purpose cleaner
library(janitor) # clean_names for snake_case
library(skimr) # skim to summarise data
library(conflicted)
conflict_prefer("filter", "dplyr")
library(kableExtra)

# the tidyverse is a metapackage that loads a suite of tools
# useful for working with and summarising data
```
<aside>
If not already installed, run `install.packages("package name")` in the console before beginning.
</aside>


Let's get some data to work with.

> todo replace with scraping from googlesheets

```{bash}
# data we want are in data-raw
ls data-raw/

```


```{r}

# import data
example_dat <- 
  read_csv("data-raw/Haddawayetal2018bufferstrips.csv") %>% 
  # clean variables names to snake_case
  clean_names()

```
<aside>
This data comes from 

> todo

</aside>



Inspecting large datasets can slow your machine down, so best to check dimensions before beginning to work with data.

```{r}
example_dat %>%
   dim()

```

With `r example_dat %>% dim() %>% pluck(1)` rows and `r example_dat %>% dim() %>% pluck(2)` columns, this is not so big that we can't inspect it. 

```{r eval=FALSE}
# inspect data 
example_dat %>% 
  View()

```

We can also use summary tools, such as `skimr::skim`.

> todo footnote about nomenclature

Running the following command in the console, we note a number of variables with only one unique (`n_unique`). 

```{r eval=FALSE}
skim(example_dat)

```

Since the `skimr::skim` summary table is a dataframe, we can identify what these variables are.

```{r}
example_dat %>% 
  # summarise
  skim() %>%
  # extract variables with only one unique character string in the column
  filter(character.n_unique == 1) %>% 
  # output name of variable from skim
  pluck("skim_variable") 

```

We can see a pattern of groupings by prefix. Now we inspect one variable: the data prefixed with `studydesign`. 

## study design

```{r}
# take a look at the study design columns
example_dat %>% 
  select(contains("studydesign"))

```

We see at least the top of the data is split with one indication in `studydesign_observational`

First, we check our assumption that there are never strings in both columns.

```{r}

# check assumption that variable is encoded across
# two columns
example_dat %>% 
  select(contains("studydesign")) %>% 
  filter(
    !( # which rows do not meet these assumptions
      
      # na in observational and manipulative string
      (is.na(studydesign_observational) & 
        studydesign_manipulative == "Manipulative") | # or
        
        # na in manipulative & observational sring
        (is.na(studydesign_manipulative) & studydesign_observational == "Observational")
    )
  )

```

The data are not _quite_ as expected. There are three levels to a condensed `studydesign` variable: observational, manipulative, and _both_ observational and manipulative.

Since we know the structure of all the data now, one way to condense these two columns would be via concatenating the variables.

```{r}
study_design <- 
  example_dat %>%
  # for this example
  select(contains("studydesign")) %>% 
  unite(
    col = "study_design",
    contains("studydesign"),
    sep = "; ",
    na.rm = TRUE
  ) 

# top of studydesign
study_design

# check the two-value columns
study_design %>% 
  filter(str_length(study_design) > str_length("Observational"))

```

## if we're happy to do this with all variables

```{r}
condense_readable <-
  function(dat, variable) {
    unite(dat, col = !!variable, 
          contains(!!variable), sep="; ", na.rm = TRUE)
  }

```

```{r}
# condense all

condensed_data <- 
  example_dat %>% 
  condense_readable("studydesign") %>% 
  condense_readable("spatialscale") %>% 
  condense_readable("measurement") %>% 
  condense_readable("farmingsystem") %>% 
  condense_readable("farmingproductionsystem") %>% 
  condense_readable("vegetationtype") %>% 
  condense_readable("striplocation") %>% 
  condense_readable("stripmanagement") 

# take a look at our condensed variables
condensed_data %>% 
  select(studydesign, spatialscale, measurement, farmingsystem, farmingproductionsystem, vegetationtype, striplocation, stripmanagement)

```

```{r}
testdat<-condensed_data %>% 
  unnest(studydesign=strsplit(studydesign, "; "))

```

