---
title: "analytical"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analytical}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(sysrevdata)
library(tidyverse)
```

## Machine-structured data

When releasing results for publication it is commmon to consider how to facilitate human interpretation through tables and visualisations. In addition, we recommend releasing data in a raw machine-readable format, with as little condensing as is realistic. Condensing data may prove to be an obstacle for extracting variables of interest by some researcher in the future.  

## Example

Here we consider an example dataset 

todo: citation

The first 12 columns are all different variables. 
```{r}
names(bufferstrips) %>% head(12)

```

However, from there the columns are mostly structured wide in groups: the first string indicates the group; the second string, the subcategory in the group, such as `spatialscale`.

```{r}
names(bufferstrips)[13:30]

```


```{r}
pivot_longer(
  data = bufferstrips,
  cols = 
)


```

