---
title: "complex"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{complex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style>
u.dotted{
  border-bottom: 1px dashed #999;
  text-decoration: none; 
}
</style>
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
# packages used in this walkthrough
library(sysrevdata)
library(tidyverse)
library(tippy)

# this avoids tidyverse conflicts with the base function filter
conflicted::conflict_prefer("filter", "dplyr")
```

## complex visualisations and analysis

In addition to a narrative synthesis, systematic reviews and maps typically require complex visualisations, such as evidence atlases. In addition, it is common for review authors to upgrade a systematic map into one or more focused systematic reviews, which may involve <u class="dotted">`r tippy("quantitative synthesis", "<span style='font-size:20px;'>e.g. meta-analysis or meta-regression</span>")`</u>. The easiest way to do this is to produce a **long** or **tidy** format, one variable per study per row. 

This is the ideal way to share data for future syntheses, as it is machine readable for multiple analyses.

In this walkthrough, we'll consider the `bufferstrips` dataset, which is wide-formatted by subcategory. 

```{r}
# spatial variables
buffer_example %>% 
  select(short_title, contains("spatial"))
```


## wide to long

We wish to convert these wide data to long-form data, dropping the `NA` elements and producing a table wherein each row represents the subcategory for one category for each study.

We'll use the collection of variables we obtained in the [creating a narrative synthesis table vignette](narrative.html). 

```{r}
buffer_variables
```

Here's one way of transforming the data to long format. 

```{r}
buffer_example_long <-
  buffer_example %>%
  # this function pivots longer
  pivot_longer(
    # see the narrative vignette for how this vector was created
    cols = contains(buffer_variables),
    # name of column we will put the column names of the wide data
    names_to = "category_type",
    # name of column we will put the values of those columns in
    values_to = "subcategory_value",
    # drop the NA values
    values_drop_na = TRUE
  ) %>%
  # I suspect there's a tidy solution to this. However, finding a regex solution 
  # was just as useful. There is a lot of string manipulation and extraction in 
  # evisynth that is not pivoting. 
  mutate(
      subcategory_type = map_chr(
    category_type,
    .f = function(x){ifelse(
    str_detect(x, "_"),
    str_match(x, "_(\\w+)") %>% pluck(2),
    NA
  )}),
    category_type = if_else(
    # extract the prefix of the column names with _
    str_detect(category_type, "_"),
    str_extract(category_type, "[a-z]*"),
    category_type
  )
  )

# newly created columns
buffer_example_long %>%
  select(short_title, category_type, subcategory_type, subcategory_value)

```

```{r eval=FALSE, include=FALSE}
usethis::use_data(buffer_example_long, overwrite=TRUE)
```


## condensed to long

Suppose, however, that we began with condensed data, as we created in the [creating a narrative synthesis table vignette](narrative.html). 

```{r}
condensed_buffer_example
```

We want to take these data and create the same long-format we have above.

```{r message=FALSE, warning=FALSE}
condensed_buffer_example %>% 
  pivot_longer(
    contains(buffer_variables),
    names_to = "category_type",
    values_to = "subcategory_value"
  ) %>% 
  separate(subcategory_value,
           # there are a maximum of 8 different subcategories
           into = letters[1:8],
           sep = "; ") %>% 
  pivot_longer(letters[1:8],
               values_to = "subcategory_value") %>%
  # get rid of nas
  filter(!is.na(subcategory_value)) %>%
  # drop redundant column
  select(-name) %>% 
  # from here is just for display
  select(short_title, category_type, subcategory_value)


```

## summarising

Now we have our data in long-form, we can perform various analyses.

We might be interested in the number of countries in the systematic review, in which case we can use the original data.

```{r}
bufferstrips %>% 
  count(study_country)
```

But to see the number of, say, observations in the farming production system, we will use our long data.

```{r}
buffer_example_long %>%
  filter(category_type == "farmingproductionsystem") %>%
  count(subcategory_value)
  
  
```

## back to wide

```{r}
back_to_wide <-
  buffer_example_long %>% 
  pivot_wider(
    id_cols = -contains("category"),
    names_from = c(category_type, subcategory_type),
    names_sep = "_",
    values_from = subcategory_value
  ) 

```



